Signatures of Foreground Sky in Power Spectrum of Redshifted Neutral Hydrogen from the Epoch of Reionization

### 2014-09-19 foreground_delay_spectrum_modeling.tex
From d78e0ca: ============
(*) Made minor change to title
(*) Added new figures
(*) Changed wording as Judd suggested
(*) Changed Tsys to 95 K and made corresponding changes to all figures
(*) Added reference(s) to eor.bib and cited them in text
(*) Introduction and nomenclature on 'baseline' yet to be reworded

%% \title{Detecting Epoch of Reionization in redshifted 21~cm: contamination in EoR window}
%% \title{Limits on the detection of the Epoch of Reionization from MWA observations of the redshifted 21~cm line}
\title{Signatures of Foreground Sky in Power Spectra of Redshifted Neutral Hydrogen from the Epoch of Reionization}

@@

Perhaps the key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}. Considerable effort has been made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,liu11,dat10,bow09,liu09}. A general model of how smooth spectrum foregrounds are scattered to higher $k$--modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum. The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}). The more aggressive alternative is to subtract a sky model and regain access to modes lost by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components most likely to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.

@@

Matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline. We confirm {\bf [JDB: is confirm the right word, seems too weak}.  Who else has published this?  Are we the first to declare it?] that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather from near the horizon.  We find that diffuse Galactic emission plays a larger role then previously thought. We quantify this contamination by separating the simulations into components based on sky region and type of foreground emission. Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.

@@

Since we are studying a redshifted H{\sc i} spectral line, $f$ is a measure of cosmological distance along the line of sight. Consequently, $\eta$ is a measure of line of sight spatial frequency mode, $k_\parallel$, given by $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of redshifted H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$. However, foreground objects emit in radio frequencies and contaminate the signals from the redshifted H{\sc i}. Due to their spatial and spectral properties, $V_\eta(\overline{\mathbf{u}},\eta)$ also contains contamination from foreground radio objects \citep{thy13,tro12,mor12,bow09}. These contaminations have to be characterized precisely in order to reduce their impact on redshifted H{\sc i} power spectrum detection sensitivity.

@@

@@

For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all antenna spacings to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$. Antenna pairs with spacings oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.

Wherever applicable, we use $H_0=70$~km~s$^{-1}$~Mpc$^{-1}$ and assume a flat $\Lambda$CDM universe with $\Omega_\textrm{M}=0.27$ and $\Omega_\Lambda=0.73$.

@@

Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=90$~K per polarization for all frequency channels for all antenna pairs throughout the course of observation. We take into account the increase in thermal uncertainty in frequency channels that results due to aforementioned bandpass correction. ({\bf Since the bandpass correction is digital, shouldn't the noise stay flat after bandpass correction? -- JDB: yes})

@@

Here we have chosen two, 112 second long sections from this night for detailed study. These were chosen to provide a selection of possible foreground and instrumental conditions. As an example of ``nominal'' observing setup we choose a zenith pointing; as an example of poor foreground conditions we choose a pointing when the field is 2 hours from zenith. This pointing has a significantly higher secondary side lobe structure and is recorded when the bright galactic center is well above the horizon.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/custom_pointings.eps}
\caption{MWA tile beam pointing directions during the course of the observation. The x-axis refers to the Local Sidereal Time (LST) in hours. The axis on the left refers to longitudes, namely, Right Ascension (RA) and Hour Angle (HA) in degrees. Negative values of RA, HA, and LST are to be interpreted as having been wrapped around by 360\arcdeg or 24 hours. The axis on the right refers to the declination of the pointing direction. The RA, HA, and declination are plotted with solid, dashed and dotted line styles respectively. A ``drift and shift'' scheme is used. The sky is allowed to drift for $\sim 30$~mins before the beamformer delay settings change in a discrete step to center the beam around RA = 00h 00m 00.00s, Dec = -30\arcdeg$\,$00\arcmin$\,$00\farcs 0. Dot--dashed vertical lines show the LST of off-zenith and zenith pointings used in our study. \label{fig:pointings}}
\end{figure}

@@

The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off12,off10}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz. These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al, in prep). MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning. It has a flux density limit of 25~mJy and a declination range of -12\arcdeg to -40\arcdeg evenly covering the field of view of the observations reported here. The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern. The calibration algorithm -- based on a forward modeling software by \citet{sul12} -- computes complex gain solutions per channel per antenna averaged to two minute intervals. The solutions are still fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation. First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass. After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variation in cable delay. Finally, we fit for an additional phase known to be caused by reflection in 150~m length cables. %possibly could add elaboration, if I knew exactly which version of the cal was done here. doesn't really matter

@@

We obtain the delay spectrum of these calibrated visibilities taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function. The spectrum is multiplied by  the additional weight of the flagged channels which occur rarely for interference and at regular spacing every 1.28~MHz where the edges of coarse channels are known to have a slight aliasing effect. In delay--space, these weights translate into a convolution by a ringing harmonic point spread function (PSF). We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12}. The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights. We limit the selection of peaks to modes inside the horizon delay, corresponding to smooth spectrum objects in the visible sky hemisphere.

@@

In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\overline{\mathbf{k}}|$. Hence, the brightest EoR signal will occur on the shortest baselines and smallest delays. Thus the region of interest for EoR studies is just beyond the horizon delay limits (solid white lines) and inside the first harmonic on either side ($|\eta| \lesssim 0.78\,\mu$s) due to the coarse channel structure. This forms a small region inside the so-called {\it EoR window}.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\caption{ Delay spectra of two 112s  Murchison Widefield Array observations of the Epoch of Reionization field at RA=0, Dec=-30. At top an {\it off--zenith}  pointing when tiles have significant side lobe response in the direction of the galactic center. At bottom a {\it zenith} pointing where side lobes are minimal and the galaxy has completely set. To get the the power spectrum each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted as a logarithmic color in order of baseline length ($x$--axis). For reference, the $k$--space equivalent of $\eta$ and $|\overline{\mathbf{x}}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively. The logarithmic color scale (shown at the right) is common to both panels. White lines mark the boundaries of the {\it foreground wedge} determined by the light travel time across each baseline. Smooth spectrum objects are physically limited to the region below this line. The various features within the wedge depend on the \emph{orientation} of the baseline, the side lobe structure of the power--pattern and the sky at that instant. \label{fig:fhd_data}}
\end{figure}

@@

The power--pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is modeled {\bf [TBD: cite Sutinjo et al (2014; submitted?)]} as a mutually-coupled 4-by-4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation. To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles. We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain \citep{bow07b}.

@@

The MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) and has non-zero response all the way to the horizon. As we have seen above, foreground objects contaminate the region inside the ``wedge'' with an amplitude that depends on the amplitude of power--pattern in that direction, with objects at the horizon being closest to the boundaries of the wedge.

The degree to which these objects leak power into spectral line of sight modes above the wedge, depends primarily on the brightness of foregrounds near the boundaries of the wedge \citep{thy13,pob13,ved12,par12}. Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting to the primary field of view.

@@

A low resolution version of the diffuse foreground model used is shown in Figure~\ref{fig:DSM}. Contours of the MWA tile power--pattern shown in Figure~\ref{fig:power_pattern} are also overlaid. Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power receptivity is significantly high ($\gtrsim 12$\%) in that direction. In the {\it zenith} pointing, the Galactic plane has almost set and the contour level of power--pattern in that direction is at least 16 times smaller.

@@

We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our compact source catalog due to their complementary survey footprints covering the entire sky, and matched flux sensitivity and angular resolution. The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec. The NVSS covers the sky with $\delta > -40$\arcdeg with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for compact objects.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and color scale are identical to those in Figure~\ref{fig:fhd_data}. The features resemble those obtained with MWA data shown in Figure~\ref{fig:fhd_data}. The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing. \label{fig:sim_data}}
\end{figure}

In order to make a quantitative comparison of delay spectra obtained with the MWA data and our simulations, we consider the following uncertainties:

{\bf Spectral index uncertainty.}  Our foreground models are derived from other higher frequency catalogs and sky maps. The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observed frequency. Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the observed frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index. From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact objects from NVSS and SUMSS catalogs. Although the model of \citet{deo08} yields a spectral index per direction on the sky, we assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from 408~MHz map of \citet{has82}. Thus, fractional errors in delay spectrum from compact foreground objects and diffuse emission are $\sim$70\% and $\sim$30\% respectively.

{\bf Thermal uncertainty.} {\bf [JDB -- this is really a model component, not an uncertainty.  We should probably move it to the earlier section.]}  In addition to intrinsic model uncertainty, delay spectra from simulations and data each have uncertainties due to thermal noise with rms, $\Delta V^\textrm{N}_\eta(|\overline{\mathbf{x}}|,\eta) \sim$ 1.4$\times 10^7$~Jy~Hz, where the superscript N stands for thermal noise. We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|\,/\,|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively. % Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.

@@

Figure~\ref{fig:data-sim-ratio} shows histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}. The median absolute deviation from these distributions correspond to $\sim$~90\% fractional difference between data and modeling on average in either case. This is in line with aforementioned uncertainties in foreground models and thermal noise in measurements. Currently we are significantly limited by unavailability of all--sky foreground models at the observed frequency of 185~MHz.

>>>>

%% \title{Detecting Epoch of Reionization in redshifted 21~cm: contamination in EoR window}
%% \title{Limits on the detection of the Epoch of Reionization from MWA observations of the redshifted 21~cm line}
\title{Signatures of Foreground Sky in Power Spectrum of Redshifted Neutral Hydrogen from the Epoch of Reionization}

@@

Perhaps the key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}. Considerable effort has been made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,liu11,dat10,bow09,liu09}. A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum. The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}). The more aggressive alternative is to subtract a sky model and regain access to modes lost by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components most likely to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.

@@

Matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline. We find that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon. We find that diffuse Galactic emission plays a larger role then previously thought. We quantify this contamination by separating the simulations into components based on type of foreground emission. Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.

@@

Since we are studying a redshifted H{\sc i} spectral line, $f$ is a measure of cosmological distance along the line of sight. Consequently, $\eta$ is a measure of line of sight spatial frequency mode, $k_\parallel$, given by $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. We use $H_0=100\,h$~km~s$^{-1}$~Mpc$^{-1}$ and assume a flat $\Lambda$CDM universe with $\Omega_\textrm{M}=0.27$ and $\Omega_\Lambda=0.73$.

Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of redshifted H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ \citep{mor04}. However, foreground objects emit in radio frequencies and contaminate the signals from the redshifted H{\sc i}. Due to their spatial and spectral properties, $V_\eta(\overline{\mathbf{u}},\eta)$ also contains contamination from foreground radio objects \citep{thy13,tro12,mor12,bow09}. These contaminations have to be characterized precisely in order to reduce their impact on redshifted H{\sc i} power spectrum detection sensitivity.

@@

Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper. $|\overline{\mathbf{x}}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis. $\eta$ and $k_\parallel$, denoting spatial scales along line of sight form the $y$--axis. Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge}, the boundaries of which are determined by antenna spacing and the light travel time across it. These boundaries are called horizon delay limits and are shown by solid lines. Spectral response of the instrument convolves the foreground wedge and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\eta$--axis. This effect has been quantified in \citet{thy13}. The width of this narrow strip is inversely proportional to operating bandwidth. Region of Fourier space excluding the {\it foreground wedge} and also the narrow strips is usually called the {\it EoR window}. These are shown in light and medium shades. However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape that results in harmonic repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s. The underlying cause of this effect will be discussed later in this paper. Hence, we have identified a region of high significance (shown in medium shade) that forms a subset of the {\it EoR window} bounded by $|\eta|\le 0.78\,\mu$s where contamination from foregrounds is expected to be low and yet the EoR H{\sc i} power spectrum signal is reasonably strong. Hereafter, we refer to this region as the {\it EoR window} in the rest of this paper. The MWA EoR collaboration actively uses this as a test region in Fourier space for diagnostics on data quality.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
\caption{Illustration of Fourier space in which delay (and power) spectra of H{\sc i} signals from the EoR are presented. The $x$--axis denoted by $|\overline{\mathbf{x}}|$ (antenna spacing in m) or $k_\perp$ (transverse wavenumber in units of $h$~Mpc$^{-1}$) are equivalent and represent transverse spatial scales. Similarly, the $y$--axis denoted by $\eta$ (delay in s or $\mu$s) or equivalently, $k_\parallel$ (line--of--sight wavenumber in units of $h$~Mpc$^{-1}$) represents spatial scales along the line of sight. The dark shaded region is often referred to as {\it foreground wedge} where smooth spectrum foregrounds reside. The boundaries of the {\it foreground wedge} (solid lines) are given by the light travel time for corresponding antenna spacings and is referred to as the horizon delay limits. In principle, region outside the wedge is expected to be relatively free of foreground contamination and is called the {\it EoR window} (light and medium shades). In practice, this contamination is also high in the narrow unshaded white regions (bound by solid and dashed lines) due to convolution by instrumental transfer function. The width of this narrow region is determined by inverse of operating bandwidth. Due to an undesirable spectral response of the instrument used in this study, namely the MWA, we identify a narrower region of high significance to EoR studies (medium shade) and hereafter refer to it as the {\it EoR window} in this paper. \label{fig:fourier-space}}
\end{figure}

@@

For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all antenna spacings to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$. Antenna spacings oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.

@@

Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=95$~K per polarization for all frequency channels for all antenna pairs throughout the course of the observation. % We take into account the increase in thermal uncertainty in frequency channels that results due to aforementioned bandpass correction. ({\bf Since the bandpass correction is digital, shouldn't the noise stay flat after bandpass correction? -- JDB: yes})

@@

Here we have chosen two, 112 second long sections from this night for detailed study. These were chosen to provide a selection of possible foreground and instrumental conditions. As an example of ``nominal'' observing setup we choose a zenith pointing; as an example of poor foreground conditions, we choose a pointing when the field is $\sim 2$~hours from zenith. This pointing has a significantly higher secondary side lobe structure and is recorded when the bright galactic center is well above the horizon.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/custom_pointings.eps}
\caption{MWA tile beam pointing directions during the course of the observation. The $x$--axis refers to the Local Sidereal Time (LST) in hours. The axis on the left refers to longitudes, namely, Right Ascension (RA) and Hour Angle (HA) in degrees. Negative values of RA, HA, and LST are to be interpreted as having been wrapped around by 360\arcdeg or 24 hours. The axis on the right refers to the declination of the pointing direction. The RA, HA, and declination are plotted with solid, dashed and dotted line styles respectively. A ``drift and shift'' scheme is used. The sky is allowed to drift for $\sim 30$~mins before the beamformer delay settings change in a discrete step to center the beam around RA = 00h 00m 00.00s, Dec = -30\arcdeg$\,$00\arcmin$\,$00\farcs 0. Dot--dashed vertical lines show the LST of off-zenith and zenith pointings used in our study. \label{fig:pointings}}
\end{figure}

@@

The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off12,off10}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz. These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al, submitted). MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning. It has a flux density limit of 25~mJy and a declination range of -12\arcdeg to -40\arcdeg evenly covering the field of view of the observations reported here. The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern. The calibration algorithm -- based on a forward modeling software by \citet{sul12} -- computes complex gain solutions per channel per antenna averaged to two minute intervals. The solutions are still fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation. First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass. After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variation in cable delay. Finally, we fit for an additional phase known to be caused by reflection in 150~m length cables. %possibly could add elaboration, if I knew exactly which version of the cal was done here. doesn't really matter

@@

We obtain the delay spectrum of these calibrated visibilities taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function. The spectrum is multiplied by  the additional weight of the flagged channels which occur rarely for interference and at regular spacing every 1.28~MHz where the edges of coarse channels are known to have a slight aliasing effect. In delay--space, these weights translate into a convolution by a ringing harmonic point spread function (PSF). We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12}. The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights. We limit the selection of peaks to modes inside the horizon delay limit, corresponding to smooth spectrum objects in the visible sky hemisphere.

@@

In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\overline{\mathbf{k}}|$. Hence, the brightest EoR signal will occur on the shortest baselines and smallest delays. Thus the region of interest for EoR studies is slightly beyond the horizon delay limits (solid white lines) and inside the first harmonic on either side ($|\eta| \lesssim 0.78\,\mu$s) due to the coarse channel structure. We call this region as the {\it EoR window} in this paper.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\caption{Delay spectra of two 112~s Murchison Widefield Array observations of the Epoch of Reionization field at RA = 0h, Dec = -30\arcdeg at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has completely set. To get the delay spectrum, each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted in order of baseline length ($x$--axis). For reference, the $k$--space equivalent of $\eta$ and $|\overline{\mathbf{x}}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively. The logarithmic color scale (shown at the right) is common to both panels. White lines mark the boundaries of the {\it foreground wedge} determined by the light travel time across each baseline. Smooth spectrum objects are physically limited to the region below this line. The various features within the {\it foreground wedge} depend on the \emph{orientation} of the antenna spacing, the secondary lobe structure of the power--pattern and the sky at that instant. \label{fig:fhd_data}}
\end{figure}

@@

The power--pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is modeled as a mutually--coupled 4--by--4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14}. To speed up simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles. We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain \citep{bow07b}.

@@

The MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) and has non--zero response all the way to the horizon. As we have seen above, foreground objects contaminate the region inside the {\it wedge} with an amplitude that depends on the amplitude of power--pattern in that direction, with objects at the horizon being closest to the boundaries of the wedge.

The degree to which these objects leak power into spectral line of sight modes above the {\it foreground wedge}, depends primarily on the brightness of foregrounds near the boundaries of the {\it wedge} \citep{thy13,pob13,ved12,par12}. Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view.

@@

A low resolution version of the diffuse foreground model used is shown in Figure~\ref{fig:DSM}. Contours of the MWA tile power--pattern shown in Figure~\ref{fig:power_pattern} are also overlaid. Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power receptivity is significantly high ($\gtrsim 12$\%) in that direction. In the {\it zenith} pointing, the Galactic plane has set and the contour level of power--pattern in that direction is at least 16 times smaller.

@@

We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our compact object catalog due to their complementary survey footprints covering the entire sky, and matched flux sensitivity and angular resolution. The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec. The NVSS covers the sky with $\delta > -40$\arcdeg with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for compact objects.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
\caption{Simulated delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), for {\it off--zenith} (top) and {\it zenith} (bottom) pointings. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and color scale are identical to those in Figure~\ref{fig:fhd_data}. The features resemble those obtained with MWA data shown in Figure~\ref{fig:fhd_data}. The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing. \label{fig:sim_data}}
\end{figure}

In order to make a quantitative comparison of delay spectra obtained with the MWA data and our simulations, we consider the uncertainty in the assumed spectral index of our foreground model. Our foreground models are derived from other higher frequency catalogs and sky maps. The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observed frequency. Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the observed frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index. From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact objects from NVSS and SUMSS catalogs. Although the model of \citet{deo08} yields a spectral index per direction on the sky, we assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from the 408~MHz map of \citet{has82}. Thus, fractional errors in delay spectrum from compact foreground objects and diffuse emission are $\sim$70\% and $\sim$30\% respectively.

In addition to intrinsic model uncertainty, delay spectra from simulations and data each have fluctuations due to thermal noise in delay spectrum with rms $\sim$ 1.4$\times 10^7$~Jy~Hz. We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|\,/\,|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively. % Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.

@@

Figure~\ref{fig:data-sim-ratio} shows histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}. The median absolute deviation from these distributions correspond to $\sim$~90\% fractional difference between data and modeling on average in either case. This is in line with aforementioned uncertainty in foreground models and thermal noise fluctuations in measurements. Currently we are significantly limited by unavailability of all--sky foreground models at the observed frequency of 185~MHz.
